Supermassive black hole and host bulge affairs \\ II. The early and late sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram

### 2015-06-01 papers/MbhMsph/mm.tex
endofmm

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, depleted core, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$, 
consistent with a dry-merging formation scenario,
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$.
At the low-mass end, the bulges of late-type (Sp) galaxies define a much steeper \emph{late-type sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by bulges with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies; 
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

\begin{itemize}
\item The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the spheroidal components of early-type (elliptical + lenticular) galaxies 
      has a slope of $1.04 \pm 0.10$ and intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
      We call this tight correlation an \emph{early-type sequence}. 
      The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the bulges of late-type (spiral) galaxies has a slope of $2-3$,
      which is less well constrained due to
      the smaller size of the subsample and, more importantly, the smaller range in spheroid stellar mass 
      ($3 \times 10^9 \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 3 \times 10^{10}$) that the subsample spans. 
      We refer to this correlation as a ``tentative'' \emph{late-type sequence}. 
      In (gas-poor) early-type galaxies, the black hole and the stellar content of the spheroidal component grow at the same pace, 
      following a linear $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      In (gas-rich) spiral galaxies, the black hole grows faster than its host bulge, 
      following a quadratic/cubic $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      For analogy with the color-magnitude diagram, 
      it may be tempting to refer to our \emph{early-type sequence} as a \emph{red sequence} 
      and to our \emph{late-type sequence} as a \emph{blue sequence}.
      Future work will explore the color of spheroids and galaxies in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.  
\item The $M_{\rm BH}/M_{\rm *,sph}$ ratio for the early-type galaxies is $0.68 \pm 0.04\%$.

was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}),

\item The $M_{\rm BH} - M_{\rm *,sph}$ log-relations for core-S\'ersic and S\'ersic spheroids have slopes consistent with each other 
      ($1.19 \pm 0.23$ and $1.48 \pm 0.20$, respectively), 
      thus these two populations do not appear to define two distinct sequences. 
\item In the $M_{\rm BH} - L_{\rm sph}$ (or $M_{\rm BH} - M_{\rm *,sph}$) diagram, for early-type galaxies, 
      we did not observe the change in slope required for consistency with the log-linear $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations.
      This issue of inconsistency remains therefore an open question. 
      It might be that we are still not probing enough low-mass black holes ($M_{\rm BH} < 10^7\rm~M_\odot$) 
      for the subsample of early-type galaxies, 
      or that the transition from  $L_{\rm sph} \propto \sigma^2$ at low luminosities to $L_{\rm sph} \propto \sigma^{(5-6)}$ at high luminosities 
      is less sharp than previously thought.
      We intend to investigate this point in our future work. 
\item Spheroids with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. 
\item The $M_{\rm BH} - L_{\rm gal}$ and $M_{\rm BH} - L_{\rm sph}$ correlations have the same level of intrinsic scatter 
      when considering early-type galaxies only. 
      Once reasonable numbers of spiral galaxies are included, 
      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$. 
\end{itemize}

>>>>

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, depleted core, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$, 
consistent with a dry-merging formation scenario,
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$.
At the low-mass end, the bulges of late-type (Sp) galaxies define a much steeper \emph{late-type sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by bulges with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies; 
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

\begin{itemize}
\item The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the spheroidal components of early-type (elliptical + lenticular) galaxies 
      has a slope of $1.04 \pm 0.10$ and intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
      We call this tight correlation an \emph{early-type sequence}. 
      The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the bulges of late-type (spiral) galaxies has a slope of $2-3$,
      which is less well constrained due to
      the smaller size of the subsample and, more importantly, the smaller range in spheroid stellar mass 
      ($3 \times 10^9 \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 3 \times 10^{10}$) that the subsample spans. 
      We refer to this correlation as a ``tentative'' \emph{late-type sequence}. 
      In (gas-poor) early-type galaxies, the black hole and the stellar content of the spheroidal component grow at the same pace, 
      following a linear $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      In (gas-rich) spiral galaxies, the black hole grows faster than its host bulge, 
      following a quadratic/cubic $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      For analogy with the color-magnitude diagram, 
      it may be tempting to refer to our \emph{early-type sequence} as a \emph{red sequence} 
      and to our \emph{late-type sequence} as a \emph{blue sequence}.
      Future work will explore the color of spheroids and galaxies in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.  
\item The $M_{\rm BH}/M_{\rm *,sph}$ ratio for the early-type galaxies is $0.68 \pm 0.04\%$. 
      This value is dramatically larger than what previously reported ($0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}).
\item The $M_{\rm BH} - M_{\rm *,sph}$ log-relations for core-S\'ersic and S\'ersic spheroids have slopes consistent with each other 
      ($1.19 \pm 0.23$ and $1.48 \pm 0.20$, respectively), 
      thus these two populations do not appear to define two distinct sequences. 
\item In the $M_{\rm BH} - L_{\rm sph}$ (or $M_{\rm BH} - M_{\rm *,sph}$) diagram, for early-type galaxies, 
      we did not observe the change in slope required for consistency with the log-linear $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations.
      This issue of inconsistency remains therefore an open question. 
      It might be that we are still not probing enough low-mass black holes ($M_{\rm BH} < 10^7\rm~M_\odot$) 
      for the subsample of early-type galaxies, 
      or that the transition from  $L_{\rm sph} \propto \sigma^2$ at low luminosities to $L_{\rm sph} \propto \sigma^{(5-6)}$ at high luminosities 
      is less sharp than previously thought.
      We intend to investigate this point in our future work. 
\item Spheroids with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. 
\item The $M_{\rm BH} - L_{\rm gal}$ and $M_{\rm BH} - L_{\rm sph}$ correlations have the same level of intrinsic scatter 
      when considering early-type galaxies only. 
      Once reasonable numbers of spiral galaxies are included, 
      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$. 
\end{itemize}
